\documentclass{../uva7310}

\begin{document}

\section{March 19}

What we've done so far:
\begin{enumerate}
    \item Measure and integral
    \item Radon Nikodym derivatives
    \item Stone-Weierstrass
\end{enumerate}

\subsection{\( L^p \), \( L^\infty \), Norms, and Banach Spaces}

Let \((X,F,\mu)\) be a measure space. Then, in addition to the \(L^1\) norm, we also have \(L^p\) and \(L^\infty \)
norms.

\begin{definition}[\(L^p\)-space]
    The space \(L^p(\mu)\) is defined to be the set of \(F\)-measurable functions, \(f:X\rightarrow\RR \), such that
    \(|f|^p\in L^1(\mu)\). Moreover, when \(1\leq p<q\leq\infty \) (and \(\mu(X)<\infty \)), we have that
    \(L^q(\mu)\subseteq L^p(\mu)\) (this will be proven later).
\end{definition}
% McCarthy: We had a discussion on ``$\frac{1}{x}$ on $[0,1]$'', but I didn't quite catch it.

\begin{definition}[\(L^\infty \)-space]
    Suppose \(\mu(X)<\infty \), then \(L^\infty(\mu)\) is the space of \(F\)-measurable functions where
    \(\text{ess}\sup |f|<\infty \). Note that the \textit{essential supremum} of \(f\) is defined as
    \(\text{ess}\sup f=\sup \{a:\exists\varepsilon>0\text{ s.t. } \mu(f^{-1}(a-\varepsilon,a+\varepsilon))>0\} \).
    This is the space of \textit{essentially bounded} functions (bounded except on a set of measure 0).
\end{definition}

\begin{definition}[Normed vector space]
    Let \(V\) be a vector space and let \(||\cdot||:V\rightarrow\RR_{\geq0}\) be a function. If \(V\) and \(||\cdot||\),
    satisfy the following properties then \((V,||\cdot||)\) is a \textit{normed vector space} (over \(\RR \)). 
    \begin{itemize}
        \item For all \(x\in V\), \(||x||=0\) implies that \(x=0\).
        \item For all \(x\in V\) and \(\lambda\in\RR \), \(||\lambda x||=|\lambda| ||x||\).
        \item For all \(x,y\in V\), \(||x+y||\leq ||x||+||y||\).
    \end{itemize}
\end{definition}

Examples of normed vector spaces:
\begin{itemize}
    \item \(L^p\) with the norm \(||f||_p={\left(\int|f|^p \right)}^{1/p}\). The triangle inequality for this norm is
    the \textit{Minkowski inequality}.
    \item \(L^\infty \) with the norm \(||f||_\infty =\text{ess}\sup |f|\).
\end{itemize}
We will prove that these are norms later.

\begin{definition}[Banach space]
    A normed vector space is \textit{Banach} if it is complete with respect to the metric \(\rho(x,y)=||x-y||\) where
    \(||\cdot||\) is the norm.
\end{definition}

We will show later that \(p>1\) implies that \(L^p\) is Banach.

Note that \(L^p\) need not be Banach for \(p\leq 1\) under usual metric. For example when \(p=1/2\),
\({(\sqrt{x}+\sqrt{y})}^2=x+y+2\sqrt{xy}\not\geq x+y\).

Exercise: Show that \({(||f-g||_p)}^p\) defines a metric on \(L^p\).

Note that all norms on \(\RR^n\) are topologically equivalent. That is if \(||\cdot||_1\) and \(||\cdot||_2\) are norms
on \(\RR^n\), then there exists a \(c>0\) such that for any \(x\), \(c^{-1}||x||_2\leq||x||_1\leq c||x||_2\). To prove
this, fix a basis \((e_i)\) of \(\RR^n\) and look at \(\min_i ||e_i||\) and \(\max_i ||e_i||\).

% TODO: make TikZ diagram
When \(X\) is the 2 point space under counting measure, consider the set of \( (x,y)\) such that
\(||(x,y)||_p={(|x|^p+|y|^p)}^{1/p}\leq 1\) (see diagram). Taking \(p=2\) will give us that the set is a disk, when
\(p=1\) the set is a diamond. As \(p\) grows larger, the set approximates a square and \(p=\infty\) gives us a square
(the norm is the max norm). When \(p<1\), the sides of the diamond start to get crushed down.

\begin{definition}[Bounded operator]
    Let \(V\) and \(W\) be vector spaces and let \(T:V\rightarrow W\) be a linear operator. Then
    \(||T||=\sup_{v\neq 0} \frac{||Tv||_W}{||v||_V}\). If \(||T||<\infty \), then \(T\) is \textit{bounded}.
\end{definition}

In finite dimensions, all operators are bounded.

Examples of operators: 
\begin{itemize}
    \item \(V=\{{(x_n)}_{n\in\NN}: {\left(\sum x_n^2\right)}^{1/2}\} \), \(T((x_n))=(nx_n)\).
    \item \(V=C^1([0,1])\), \(W=C([0,1])\), \(T=\frac{d}{dx}:C^1\rightarrow C\) which is a dense domain. Here \(T\) is
    not bounded, however this is coming from the fact that its not always defined.
\end{itemize}

To get a truly unbounded operator, need axiom of choice (e.g.\ it might as well be a myth). Take bases of Banach spaces
\(V\) and \(W\), \((e_\alpha)\) and \((f_\beta)\). Take a subcollection \((e_i)\subseteq (e_\alpha)\) and then map
\(e_i\mapsto ie_i\).

An operator is bounded iff it is continuous since \(||T(v+h)-T(v)||=||T(h)||\leq ||T||\cdot ||h||\).

Given \(V\), take \(V^*\) to be the dual space (the space of bounded functionals on \(V\)) and give it the operator
norm.

\subsection{Convexity}

\begin{definition}[Convex set]
    A set \(A\subseteq \RR^d\) is \textit{convex} if for all \(x,y\in A\) and \(\alpha\in[0,1]\) we have that
    \(\alpha x+(1-\alpha)y\in A\).
\end{definition}

Exercise: All convex sets are Borel measureable. Sketch: Proceed via induction on \(d\). It is true in \(\RR^1\). Slice
\(A\) into \(A_{x_1}\subseteq\RR^{d-1}\) slices for fixed \(x_1\).

\begin{definition}[Convex function]
    A function \(f\) is \textit{convex} if its epigraph, \( \{(x,y):y\geq f(x)\} \) is convex. That is, if
    \(f(\alpha x+(1-\alpha) y)\leq \alpha f(x)+(1-\alpha) f(y)\) for all \(x,y\) and \(\alpha\in[0,1]\). Alternatively,
    if for all \(x,y,z\) with \(x<y<z\) we have that \(\frac{f(z)-f(y)}{z-y}\geq \frac{f(y)-f(x)}{y-x}\).
\end{definition}

\begin{proposition}
    Any convex function is continuous.
\end{proposition}
\begin{proof}
    Suppose \(f:\RR\rightarrow\RR \) and \(x_2<x<y<x_1\). Then 
    \[
        \frac{f(x_1)-f(y)}{x_1-y}\geq\frac{f(y)-f(x)}{y-x}\geq \frac{f(y)-f(x_2)}{y-x_2}
    \]
    and slopes are bounded and increase as \(x\) increases. Ergo, this converges as \(x\to y\) and thus \(f\) is
    continuous and has directional derivatives. If \(f:\RR^d\rightarrow\RR\) then \(g(t)=f(x+tv)\) is convex in \(t\)
    and is thus has directional derivatives (since it is a convex function from \(\RR \) to \(\RR \)).
    Ergo, \(D_v^\pm = \lim_{t\to 0\pm} \frac{f(x+tv)-f(x)}{t}\) and \(f\) is continuous.
\end{proof}
\begin{corollary}
    Since convex functions are continuous, they are Borel measureable.
\end{corollary}

Recall that if \(X,F,\mu \) is a measure space and \(f:X\rightarrow \RR \) is measurable we can push forward \(\mu \)
to \(\RR \) with \(f_*\mu(E)=\mu(f^{-1}(E)) \).

\begin{thm}[Jensen's Inequality]
If \(f\in L^1(X,F,\mu)\) with \(\mu(X)<\infty \) and \(\varphi:\RR\rightarrow\RR \) is convex then
    \[
        \varphi\left(\frac{1}{\mu(X)}\int fd\mu\right)\leq \frac{1}{\mu(X)}\int \varphi(f(x))\mu(dx).
    \]
\end{thm}
\begin{proof}
    Let \(E=\frac{1}{\mu(X)}\int fd\mu \) and
    \(\beta=D^-\varphi(E)=\lim_{\varepsilon\to 0-}\frac{\varphi(E+\varepsilon)-varphi(E)}{\varepsilon}\leq D^+\varphi(E)\).
    Then \(\varphi(s)-\varphi(E)\geq \beta(s-E)\). Ergo, \(\varphi(f(x))-\varphi(E)\geq\beta(f(x)-E)\). Thus,
    \(\frac{1}{\mu(X)}\int d\mu(x)\geq 0\).
\end{proof}

\begin{corollary}
    \begin{enumerate}
        \item Young inequality: If \(\varphi(t)=e^t\) and \(\mu(X)=1\) then \(e^{\int f}\leq \int e^f\).
        \item AMGM inequality: Take \(X=\{1,\ldots, n\} \) and let \(\alpha_1,\ldots,\alpha_n>0\) with \(\sum a_i=1\).
        Let \(\mu(\{i\})=\alpha_i\). Then \(y_i=e^{f(i)}\geq 0\) and \(\prod y_i^{\alpha_i} \leq \sum \alpha_i y_i\).
    \end{enumerate}
\end{corollary}

\subsection{What we covered today}

Normed spaces, Banach spaces, bounded operators

Convexity

Jensen's Inequality

\(L^p\), \(L^\infty \)

\subsection{Goals for next time}

\(L^q\subseteq L^p\) when \(q>p>1\), Banach, \((L^p)^*\).

Jensen, (H\"older, Minkowski).

Suppose \(f\in L^1\mapsto T(f)\in\RR \). Then \(T(f)=\int f d\mu \) is an example of a functional. We can also have
\(T_A(f)=\int 1_A f\). Can take \(T_g(f)=\int f\bar{g} d\mu \) for \(g\in L^\infty \).

We will see that \({(L^1)}^*=L^\infty \), but \({(L^\infty)}^*\neq L^1\).

\section{March 21}

Assume $\mu(X)$ finite.

Recall
\begin{itemize}
    \item Convex
    \item Jensen ($\mu(X)=1$) $\phi(\int fd\mu)\leq \int\phi(f(x))d\mu(x)$.
    \item Young
\end{itemize}

If $\phi:\RR\rightarrow\RR$ convex and $f$ is $L^1$, is $\phi\circ f\in L^1$? If $\phi\circ f\notin L^1$ then $\int \phi\circ f=+\infty$.

H\"older inequality

Def: $||f||_p=\left(\int|f|^p\right)^{1/p}$
Def: Conjugate, $p\to p'$ s.t. that $1/p+1/p'=1$ then $p'=p/(p-1)$.

Want to show $||f||_p$ is a norm in $L^p$.

H\"older inequality: If $f\in L^p$ and $g\in L^{p'}$, then $||fg||_1\leq ||f||_p||g||_{p'}$.

Cor (Cauchy-Schwarz): $p=2$, then $\int fg \leq \sqrt{\int f^2\int g^2}$ and $||x_1x_2+y_1y_2|\leq \sqrt{(x_1^2+x_2^2)(y_1^2+y_2)^2}$.

Cor: $p=1$, $p'=\infty$, then $\int |fg|\leq \int |f|\text{ess}\sup |g|$.

Proof of H\"older

Assume $p\in(1,\infty)$ and $f,g\geq 0$ with $||f||_p\neq 0$ and $||g||_{p'}\neq 0$. Define $F=f/||f||_p$ and $G=g/||g||_{p'}$.
Then we claim that $FG\leq \alpha F^p+(1-\alpha) G^{p'}$. Consider $F^{p\alpha}G^{p'(1-\alpha)}$ with $\alpha=1/p$. Then
AM-GM inequality yields $FG=F^{p/p}G^{p'(1-1/p)}\leq \frac{1}{p}F^p+(1-\frac{1}{p}) G^{p'}$.
Moreover $\int F^p=\int G^{p'}=1$ and so $\int FG\leq 1$. QED.

Minkowski inequality: $||f+g||_p\leq ||f||_p+||g||_p$.

Proof: We can assume that $p\in(1,\infty)$. Then $t\mapsto t^p$ is convex for such $p$. Therefore, $((f+g)/2)^p\leq f^p/2+g^p/2$
which implies $f+g\in L^p$. Furthermore, $(f+g)^p=f(f+g)^{p-1}+g(f+g)^{p-1}$ and $(f+g)^{(p-1)p'}=(f+g)^p\in L^1$. Thus we can
use H\"older to get that $\int |f(f+g)^{p-1}|\leq ||f||_p \cdot \left(\int|f+g|^p\right)^{1/p'}$. Similarly,
$\int|g(f+g)^{p-1}\leq ||g||_p\left(|f+g|^p\right)^{1/p'}$. Thus $\int |f+g|^p\leq ...$. Almost QED

Cor: $L^p$ is a normed vector space. (Above proof showed vector space followed by norm).

What about when $p<1$? $L^p$ is still a vector space, but $||\alpha||_p=|\alpha|||f||_p$ and $||f||_p=0\Rightarrow f=0$ a.e..
However, no triangle inequality. But $||f+g||_p\leq 2^{1/p-1} (||f||_p+||g||_p)$. This is a \textit{quasi-norm}.
We still have a metric on this space. This is left as an exercise.

Thm: $L^p$ is Banach for $1\leq p< \infty$.

Pf: Left to show that $L^p$ is complete with respect to norm. Suppose $(f_n)_{n\in\NN}$ is Cauchy in $L^p$. Ergo,
$(f_n)$ is Cauchy in $\mu$ by Chebyshev ($\varepsilon \mu(\{x:|g(x)|>\varepsilon\})^{1/p}\leq ||g|_p$ for all $\varepsilon>0$).
This implies that there exists a subsequence $(f_{n_j})$ that converges to $f$ a.e.. Want $f_n\to f$ in $L^p$.
We know that for $n,m>N$ we have $||f_n-f_m||_p<\varepsilon$. Fatou implies that when $m>N$,
$||f-f_m||_p \leq \liminf_{j:n_j>N} ||f_{n_j}-f_m||_p<\varepsilon$. Thus $f_m\to f$ in $L^p$. QED.

Inclusion: If $\mu(X)<\infty$ then $L^p\subseteq L^q$ for all $\infty\geq q>p\geq 1$.

Note: If $X=\RR$, $\mu$ is Lebesgue. Take $f=1/x\; 1_{x>1}$ then $f\in L^{1+\delta}$ and $f\notin L^1$.

Pf of inclusion: Take $r=q/p>1$, $r'$ such that $1/r+1/r'=1$.
Write $||f||_p=||f\cdot 1||_p=\left(\int |f|^p\cdot 1\right)^{1/p}\leq \left[  \left(\int|f|^{pq/p}\right)^{p/q}\mu(X)^{1/r'} \right]^{1/p}=k(X,p,q)||f||_q$
where $k$ is a constant dependent on $X$, $p$, and $q$. QED.

Riesz representation theorems:

Def: If $V$ is a Banach space, $V^*$ is the space of linear functionals $\phi:V\rightarrow\RR$ with $||\phi||_{V^*}=\sup_{x\neq 0}\frac{|\phi(x)|}{||x||}$.

What is $(L^p)^*$?

Hahn-Banach Thm: If $V$ is a Banach space and $U$ is a linear subspace over $\RR$ and $\phi:U\rightarrow\RR$ a linear functional
where $\phi(x)\leq p(x)$ for all $x\in U$ with $p$ convex, then there exists a functional $\psi:V\rightarrow\RR$ such that
$\psi_U=\phi$ and $\psi(x)\leq p(x)$ for all $x\in V$. 

Thm: $(L^p)^*=L^{p'}$ for $1\leq p<\infty$ and $(L^\infty)^*\neq L^1$.

IntuitionL Let $\lambda(f)$ be a functional with $f\in L^p$. Identify it with $\int fgd\mu$ where $g$ a fixed function in $L^{p'}$.

Pf: 
We start by showing $(L^\infty)^*$ is not $L^1$. Take $L^\infty([0,1])=V$ (with Lebesgue measure) and $U$ to be the set of functions continuous on $[0,1]$
except maybe at $x=1/2$. Define $\phi(f)=\lim_{x\to 1/2 +} f(x)-\lim_{x\to 1/2 -}f(x)$. By Hahn-Banach, this extends to $\psi$ on $V$.
Assume there exists an $h$ such that $\psi(f)=\int_0^1 h\cdot f d x$. We can take an open set $E\subseteq [0,1]$ with $f_n\to 1_E$ ($f_n\in C([0,1])$).
Then $\psi(f_n)=0$ but $\int h f_n\to \int h 1_E=0$ by DCT. Thus $h=0$ and $\psi=0$. But $\psi$ cannot be (see any function with a jump at 1/2).
Ergo, $(L^\infty)^*\neq L^1$.

We now show that $(L^p)^*=L^{p'}$. Let $\Lambda(f)$ be a functional. We want to find $g\in L^{p'}$ such that $\Lambda(f)=\int fg$.
Define a signed measure $\nu(A)=\Lambda (1_A)$. Without loss of generality take $\mu(X)=1$ and $||\Lambda||=1$ (normalize).
Thus, $|\nu(A)|\leq ||1_A||_p=\mu(A)^{1/p}$. Moreover $\nu$ is $\sigma$-additive. So, $E=E_1\cup E_2\cup\ldots$ and
$1_E$ is the $L^p$ limit of $1_{E_1\cup \dots\cup E_n}$ and $\Lambda$ is continuous. Thus $\nu$ is indeed a signed measure
that is absolutely continuous with respect to $\mu$. Ergo, there exists a Radon-Nikodym derivative $h$ so that
$\nu(A)=\int_A h d\mu$. This further implies that for all $f\in L^p$, $\Lambda(f)=\int hf d\mu\leq ||f||_p$. We still need $h\in L^{p'}$.
We already know $h\in L^1$. The following lemma completes the proof.

Lem: $\int |hf|\leq c||f||_p$ for all $f\in L^p$ then $h\in L^{p'}$.

Pf: WLOG: assume $c=1$. Suppose $h_n\to h$ monotonically with all $h_n$'s bounded. Then $\int h_n^{p'}=\int h_n^{p'-1}h_n\leq ||h_n^{p'-1}||_p$ by
assumption. Then $(p'-1)p=p'$ so $\int h_n^{p'}\leq ||h_n||_{p'}^{p'/p}$. Thus $||h_n||_{p'}\leq ||h_n||_{p'^{1/p}}$ and
$||h_n||_{p'}\leq 1$. Ergo, $||h||_{p'}\leq 1$. QED.
\end{document}